Some examples in extended logic counting classes

### 2016-12-23 LICS/appendix.tex

For a given signature $\R$, we define $\ostr[\R]^*$ as $$\ostr[\R]^* = \{(\A,v,V) \mid \A\in\ostr[\R]\text{, $v$ ($V$) is a first-order (second-order) assignment for $\A$}  \}.$$
The {\em conditional count} symbol $(\varphi \mapsto \alpha)$ is defined as $(\neg\varphi + (\varphi\cdot\alpha))$ for given $\so$ formula $\varphi$ and $\qso$ formula $\alpha$. Note that for each $(\A,v,V) \in \ostr[\R]^*$, 
$$
\sem{(\varphi \mapsto \alpha)}(\A,v,V) = 
\begin{cases}
\sem{\alpha}(\A,v,V) &\text{if } (\A,v,V)\models\varphi,\\
0 &\text{otherwise}.
\end{cases}
$$
We will use the symbol $<$ also to denote the lexicographic order over same-sized tuples. If $\bar{x} = (x_1,\ldots,x_m)$ and $\bar{y} = (y_1,\ldots,y_m)$ are tuples of first-order variables, we denote $\bar{x} < \bar{y}$ for the formula $\bigvee_{i = 1}^m[\bigwedge_{j = 1}^{i-1}x_j = y_j \wedge x_i < y_i]$. Similarly, we use $=$ to denote equality between tuples, as $\bar{x} = \bar{y}$ denotes $\bigwedge_{i = 1}^m(x_i = y_i)$.

If $\bar{x} = (x_1,\ldots,x_m)$ ($\bar{X} = (X_1,\ldots,X_m)$) is a tuple of first-order (second-order) variables, we denote $\sa{\bar{x}}\alpha$ for $\sa{x_1}\cdots\sa{x_m}\alpha$ and $\sa{\bar{X}}\alpha$ for $\sa{X_1}\cdots\sa{X_m}\alpha$ for each $\qso$ formula $\alpha$.
\subsection{Proof of Theorem \ref{no-mult}}

@@

First, for a given signature $\R$ and logic fragment $\LL$ we define a {\em valid form in $\LL$} (or just {\em valid form}). A formula $\alpha$ is in valid form in $\LL$ if (1) it is equal to $\varphi$, $1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{x}}\sa{\bar{X}}\varphi$ or $\sa{\bar{x}}\sa{\bar{X}}1$ for some $\LL$ formula $\varphi$ and some tuple of first-order (second-order) variables $\bar{x}$ ($\bar{X}$), or (2) if it is equal to $(\beta\add\gamma)$ where $\beta$ and $\gamma$ are formulas in valid form. Given a $\eqso(\loge{0})$ formula $\alpha$ over a signature $\R$ we will recursively construct a formula $\tau(\alpha)$ in valid form such that $\alpha(\enc(\A)) = \tau(\alpha)(\enc(\A))$ for each $\A\in\ostr[\R]$. First, each natural constant $s$ in $\alpha$ is replaced by $(1\add 1\add\cdots\add 1)$ ($s$ times). If $\alpha = 1$ or $\alpha = \varphi \in \loge{0}$, then $\tau(\alpha) = \alpha$. We assume that every sub-formula of $\alpha$ is in valid form. If $\alpha = (\beta \add \gamma)$ then $\tau(\alpha) = \alpha$. If $\alpha = \sa{x}\beta$ or $\alpha = \sa{X}\beta$ this case separates in more cases: If $\alpha = \sa{x}\beta$ and $\beta$ is in one of the valid forms in (1), then $\tau(\alpha) = \alpha$. If $\alpha = \sa{X}\beta$ and $\beta$ is in the valid form $\sa{\bar{x}}\gamma$ then $\tau(\alpha) = \sa{\bar{x}}\sa{X}\gamma$; if $\beta$ it is in one of the remaining valid forms in (1), then $\tau(\alpha) = \alpha$. If $\alpha = \sa{x}\beta$ ($\alpha = \sa{X}\beta$) and $\beta = (\gamma_1\add\gamma_2)$, then $\tau(\alpha) = \tau(\sa{x}\gamma_1) \add \tau(\sa{x}\gamma_2)$ ($\tau(\alpha) = \tau(\sa{X}\gamma_1) \add \tau(\sa{X}\gamma_2)$). This covers all possible cases for $\alpha$. Let $\alpha$ be a $\eqso(\loge{0})$ formula over a signature $\R$. Consider the following procedure that receives $\enc(\A)$ as input, where $\A\in\ostr[\R]$ with domain $A$. Let $\tau(\alpha) = \beta_1\add\cdots\add\beta_m$. For each $\beta_i$, if $\beta_i = 1$ or $\beta_i = \varphi$, then $\sem{\beta_i}(\A)$ can be computed in polynomial time. If $\beta_i = \sa{\bar{x}}1$ then $\sem{\beta_i}(\A) = \vert A \vert^{\length(\bar{x})}$, if $\beta_i = \sa{\bar{X}}1$ then $\sem{\beta_i}(\A) = 2^{\Sigma_{X\in\bar{X}}\vert A \vert^{\arity(X)}}$, if $\beta_i = \sa{\bar{x}}\sa{\bar{X}}1$ then $\sem{\beta_i}(\A) = \vert A \vert^{\length(\bar{x})}2^{\Sigma_{X\in\bar{X}}\vert A \vert^{\arity(X)}}$, all of which can be computed in polynomial time. Otherwise, $\beta_i$ is $\sa{\bar{x}}\varphi$ or $\sa{\bar{X}}\sa{\bar{x}}\varphi$, and then $\sem{\beta_i}(\A) = f_{\varphi}(\enc(\A))$, and $f_{\varphi}\in \#\loge{0}$ which implies that $f_{\varphi}(\enc(\A))$ can be computed in polynomial time \cite{DBLP:journals/jcss/SalujaST95}.

@@

Consider an $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$ where $\z = (z_1,\dots,z_d)$ and $\X = (X_1,\dots,X_r)$. There are three possibilities regarding the size of the tuples of free variables $\z$ and $\X$: (1) $d>0$ and $r=0$ (2) $d=0$ and $r>0$ (3) $d,r>0$. This separates the proof in three cases:
\begin{enumerate}
	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$. That is,
	\begin{eqnarray*}
		f(\A) &=& \mid \{ \langle\e\rangle \mid \A \models \exists \x \ \varphi(\x,\e) \} \mid,
	\end{eqnarray*}
	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\e \in A^d$. Our goal here is to eliminate the lexicographically smallest sequence of variables, which can be done easily. First, let $\y = (y_1,\dots,y_k)$, $\y\,^\prime = (y_1^\prime,\dots,y_k^\prime)$ and
	\begin{eqnarray*}
		\varphi_{k,<}(\y\,^\prime,\y) &=& \bigvee_{i = 1}^k \left( \bigwedge_{j=1}^{i-1} y_j^\prime = y_j \wedge y_i^\prime < y_i \right).
	\end{eqnarray*}
	This formula is true if $\y\,^\prime$ is lexicographically smaller than $\y$. Now, let $f'$ be defined by
	\begin{eqnarray*}
		\varphi^\prime(\x,\z) &=& \varphi(\x,\z) \wedge \exists \z\,^\prime (\varphi(\x,\z\,^\prime) \wedge \varphi_{d,<}(\z\,^\prime,\z ) ).
	\end{eqnarray*}
	If $f(\A)>0$, then $f'(\A)$ will count exactly one element less than $f(\A)$. Otherwise, if $f(\A)=0$, then $\A \not\models\exists \x\,\varphi(\x,\e)$ for every tuple $\e$ of elements in $A$, so $\A \not\models\exists \x\,\varphi^\prime(\x,\e)$ for every $\e$ and, therefore, $f'(\A)=0$. Hence, $f' = f\dotminus 1,$ from which we conclude that $f\dotminus 1\in\E{1}.$

	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,\dots,x_d)$ and $\X = (X_1,\dots,X_r)$. That is,
	\begin{eqnarray*}
		f(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \exists \x \ \varphi(\x,\P) \} \mid \label{f1},
	\end{eqnarray*}
	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\P = (P_1,\ldots,P_r)$ and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$. For the time being, suppose that
	\begin{eqnarray}
	\varphi(\x,\X) &=& \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)
	\end{eqnarray}
	where $n$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, $\y$ is a $p$-tuple of variables in $\x$, $\varphi^{-}(\X,\y)$ is a conjunction of negated predicates in $\X$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\beta(\x)$ is an FO-formula over $\L$ which mentions all variables in $\x$. Note that $\theta(\x)$ also mentions all variables in $\x$. We also assume that $(\x_1,\dots,\x_n,\y) = \x$. As an example, the following formula is of this form:
	\begin{multline*}
	\varphi(\x,\X) =  X_1(x_1,x_2) \wedge X_3(x_3) \wedge X_2(x_4,x_5) \wedge X_3(x_6) \wedge \neg X_1(x_7,x_8) \wedge \\ (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 \wedge x_2 = x_8 \wedge x_2 = x_5 \wedge x_8 < x_6 \wedge x_6 = x_7 ) \wedge \\ \forall z\big( S_1(x_1,z,x_2) \wedge x_3 = x_3 \wedge x_4 = x_4 \wedge x_5 = x_5 \wedge x_6 = x_6 \wedge x_7 = x_7 \wedge x_8 = x_8 \big),
	\end{multline*}
	where $\x = (x_1,x_2,x_3,x_4,x_5,x_6,x_7,x_8)$ and $\X = (X_1,X_2,X_3)$. Here, $n = 4$, $\lambda(1) = 1$, $\lambda(2) = \lambda(4) = 3$ and $\lambda(3) = 2$, $\x_1 = (x_1,x_2)$, $\x_2 = (x_3)$, $\x_3 = (x_4,x_5)$, $\x_4 = (x_6)$ and $\y = (x_7,x_8)$. Moreover, $\varphi^{-}(\X,\y) = \neg X_1(x_7,x_8)$, $\theta(\x) = (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 \wedge x_2 = x_8 \wedge x_2 = x_5 \wedge x_8 < x_6 \wedge x_6 = x_7 )$, which defines a total order on the partition of $\x$ $\{\{x_1,x_3,x_4\},\{x_2,x_5,x_8\},\{x_6,x_7\}\}$, and $\beta(\x) = \forall z\big( S_1(x_1,z,x_2) \wedge x_3 = x_3 \wedge x_4 = x_4 \wedge x_5 = x_5 \wedge x_6 = x_6 \wedge x_7 = x_7 \wedge x_8 = x_8 \big)$.

@@

	\alpha_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha(\v_1,\dots,\v_n,\w)\wedge\bigvee_{i=1}^n(\u_i\neq\v_i)\big)\to \varphi_{m,<}((\u_1,\dots,\u_n),(\v_1,\dots,\v_n))\Big)\Big],

>>>>

For a given signature $\R$, we define $\ostr[\R]^*$ as $$\ostr[\R]^* = \{(\A,v,V) \mid \A\in\ostr[\R]\text{, $v$ ($V$) is a first-order (second-order) assignment for $\A$}  \}.$$
The {\em conditional count} symbol $(\varphi \mapsto \alpha)$ is defined as $(\neg\varphi + (\varphi\cdot\alpha))$ for given $\so$ formula $\varphi$ and $\qso$ formula $\alpha$. Note that for each $(\A,v,V) \in \ostr[\R]^*$, 
$$
\sem{(\varphi \mapsto \alpha)}(\A,v,V) = 
\begin{cases}
\sem{\alpha}(\A,v,V) &\text{if } (\A,v,V)\models\varphi,\\
0 &\text{otherwise}.
\end{cases}
$$
We will use the symbol $<$ also to denote the lexicographic order over same-sized tuples. If $\bar{x} = (x_1,\ldots,x_m)$ and $\bar{y} = (y_1,\ldots,y_m)$ are tuples of first-order variables, we denote $\bar{x} < \bar{y}$ for the formula $\bigvee_{i = 1}^m[\bigwedge_{j = 1}^{i-1}x_j = y_j \wedge x_i < y_i]$. Similarly, we use $=$ to denote equality between tuples, as $\bar{x} = \bar{y}$ denotes $\bigwedge_{i = 1}^m(x_i = y_i)$, and also $\bar{x}\leq\bar{y}$ denotes $\bar{x} < \bar{y} \vee \bar{x} = \bar{y}$. We also denote $\min(\bar{x}) := \forall\bar{y}(\bar{y} < \bar{x}\vee \bar{y} = \bar{x})$.

If $\bar{x} = (x_1,\ldots,x_m)$ ($\bar{X} = (X_1,\ldots,X_m)$) is a tuple of first-order (second-order) variables, we denote $\sa{\bar{x}}\alpha$ for $\sa{x_1}\cdots\sa{x_m}\alpha$ and $\sa{\bar{X}}\alpha$ for $\sa{X_1}\cdots\sa{X_m}\alpha$ for each $\qso$ formula $\alpha$. We also denote $\length(\bar{x}) = m$ ($\length(\bar{X}) = m$).
\subsection{Proof of Theorem \ref{no-mult}}

@@

First, for a given signature $\R$ and logic fragment $\LL$ we define a {\em valid form in $\LL$} (or just {\em valid form}). A formula $\alpha$ is in valid form in $\LL$ if (1) it is equal to $\varphi$, $1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{x}}\sa{\bar{X}}\varphi$ or $\sa{\bar{x}}\sa{\bar{X}}1$ for some $\LL$ formula $\varphi$ and some tuple of first-order (second-order) variables $\bar{x}$ ($\bar{X}$), or (2) if it is equal to $(\beta\add\gamma)$ where $\beta$ and $\gamma$ are formulas in valid form. Given a $\eqso(\loge{0})$ formula $\alpha$ over a signature $\R$ we will recursively construct a formula $\tau(\alpha)$ in valid form such that $\alpha(\enc(\A)) = \tau(\alpha)(\enc(\A))$ for each $\A\in\ostr[\R]$.

First, each natural constant $s$ in $\alpha$ is replaced by $(1\add 1\add\cdots\add 1)$ ($s$ times). If $\alpha = 1$ or $\alpha = \varphi \in \loge{0}$, then $\tau(\alpha) = \alpha$. We assume that every sub-formula of $\alpha$ is in valid form. If $\alpha = (\beta \add \gamma)$ then $\tau(\alpha) = \alpha$. If $\alpha = \sa{x}\beta$ or $\alpha = \sa{X}\beta$ this case separates in more cases: If $\alpha = \sa{x}\beta$ and $\beta$ is in one of the valid forms in (1), then $\tau(\alpha) = \alpha$. If $\alpha = \sa{X}\beta$ and $\beta$ is in the valid form $\sa{\bar{x}}\gamma$ then $\tau(\alpha) = \sa{\bar{x}}\sa{X}\gamma$; if $\beta$ it is in one of the remaining valid forms in (1), then $\tau(\alpha) = \alpha$. If $\alpha = \sa{x}\beta$ ($\alpha = \sa{X}\beta$) and $\beta = (\gamma_1\add\gamma_2)$, then $\tau(\alpha) = \tau(\sa{x}\gamma_1) \add \tau(\sa{x}\gamma_2)$ ($\tau(\alpha) = \tau(\sa{X}\gamma_1) \add \tau(\sa{X}\gamma_2)$). This covers all possible cases for $\alpha$. Furthermore, suppose that $\LL$ is closed under conjunction and suppose $\alpha = (\beta\mult\gamma)$. If $\beta = (\delta_1\add\delta_2)$ (wlog), then $\tau(\alpha) = \tau(\delta_1\mult\gamma)\add\tau(\delta_2\mult\gamma)$. If $\beta = \sa{\bar{x}}\delta$, then $\tau(\alpha) = \tau(\sa{\bar{x}}\tau(\delta\mult\gamma))$. If $\beta = 1$, then $\tau(\alpha) = \gamma$. If $\beta = \varphi$ and $\gamma = \psi$, then $\tau(\alpha) = \varphi\wedge\psi$. Let $\alpha$ be a $\eqso(\loge{0})$ formula over a signature $\R$. Consider the following procedure that receives $\enc(\A)$ as input, where $\A\in\ostr[\R]$ with domain $A$. Let $\tau(\alpha) = \beta_1\add\cdots\add\beta_m$. For each $\beta_i$, if $\beta_i = 1$ or $\beta_i = \varphi$, then $\sem{\beta_i}(\A)$ can be computed in polynomial time. If $\beta_i = \sa{\bar{x}}1$ then $\sem{\beta_i}(\A) = \vert A \vert^{\length(\bar{x})}$, if $\beta_i = \sa{\bar{X}}1$ then $\sem{\beta_i}(\A) = 2^{\Sigma_{X\in\bar{X}}\vert A \vert^{\arity(X)}}$, if $\beta_i = \sa{\bar{x}}\sa{\bar{X}}1$ then $\sem{\beta_i}(\A) = \vert A \vert^{\length(\bar{x})}2^{\Sigma_{X\in\bar{X}}\vert A \vert^{\arity(X)}}$, all of which can be computed in polynomial time. Otherwise, $\beta_i$ is $\sa{\bar{x}}\varphi$ or $\sa{\bar{X}}\sa{\bar{x}}\varphi$, and then $\sem{\beta_i}(\A) = f_{\varphi}(\enc(\A))$, and $f_{\varphi}\in \#\loge{0}$ which implies that $f_{\varphi}(\enc(\A))$ can be computed in polynomial time \cite{DBLP:journals/jcss/SalujaST95}.

@@

Let $\alpha$ be a $\eqso(\logex{1})$ formula over a signature $\R$. We will define a $\eqso(\logex{1})$ formula $\kappa(\alpha)$ such that for each finite structure $A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$. Let $\tau(\alpha) = \beta_1\add\cdots\add\beta_n$ be defined as in \ref{eqso-sigma-zero-in-fp}, that is, each $\beta_i$ is equal to $1$, $\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{x}}\sa{\bar{X}}1$ or $\sa{\bar{x}}\sa{\bar{X}}\varphi$, where $\varphi$ is a formula in $\logex{1}$ and $\bar{x}$ ($\bar{X}$) is a tuple of first-order (second-order) variables. We further assume that each $\beta_i$ is either $\sa{\bar{x}}\varphi$ or $\sa{\bar{x}}\sa{\bar{X}}\varphi$, since each $1$, $\varphi$ and $\sa{\bar{X}}\varphi$ can be replaced by the formulas $\sa{x}\min(x)$, $\sa{x}(\varphi\wedge\min(x))$ and $\sa{x}\sa{\bar{X}}(\varphi\wedge\min(x))$ respectively, while maintaining the value of $\sem{\alpha}(\A)$.

The intuition behind our reasoning is separated in three points: (1) For each $\beta_i$ of the form $\sa{\bar{x}}\varphi$, the formula $\kappa(\beta_i)$ will count every tuple $\bar{x}$ that satisfies $\varphi$ except for the lexicographically smallest one. (2) For each $\beta_i$ of the form $\sa{\bar{x}}\sa{\bar{X}}\varphi$, the formula $\kappa(\beta_i)$ will isolate the smallest $\bar{X}$ that satisfies $\varphi$, and exclude the lexicographically smallest tuple $\bar{x}$ that satisfies $\varphi(\bar{X})$. And (3) if $\alpha = (\beta\add\sa{\bar{x}}\varphi)$ or $\alpha = (\beta\add\sa{\bar{x}}\sa{\bar{X}}\varphi)$, the formula $\kappa(\alpha)$ will exclude the lexicographically smallest tuple that satisfies $\varphi$ if and only if $\beta$ is equal to 0.

For each $\alpha$ in $\eqso(\logex{1})$ such that $\alpha = \sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})$ for some quantifier-free formula $\varphi$, we define $\kappa(\alpha) = \sa{\bar{x}}\exists\bar{y}[\varphi(\bar{x},\bar{y})\wedge\exists\bar{z}(\varphi(\bar{z},\bar{y})\wedge\bar{z} < \bar{x})]$, which is in $\eqso(\logex{1})$ and fulfils the desired condition.

For each $\alpha$ in $\eqso(\logex{1})$ such that $\alpha = \sa{\bar{x}}\sa{\bar{X}}\exists\bar{y}\,\varphi(\bar{x},\bar{X},\bar{y})$ for some quantifier-free formula $\varphi$, we define $\kappa(\alpha)$ procedurally as follows: Let $\bar{x} = (x_1,\ldots,x_{\length(\bar{x})})$ and $\bar{X} = (X_1,\ldots,X_{\length(\bar{X})})$. We suppose that $\varphi$ is in a DNF form that leaves formulas that do not mention $\bar{X}$ as is. If it is not, we convert $\varphi$ to this form with a standard transformation algorithm. Let $\varphi(\bar{x},\bar{X},\bar{y}) = \bigvee_{i = 1}^m\varphi_i(\bar{x},\bar{X},\bar{y})$ where each $\varphi_i$ has the form:
$$
\varphi_i(\bar{x},\bar{X},\bar{y}) = (\text{conjunction of $X_i$'s}) \wedge (\text{conjunction of $\neg X_i$'s})  \wedge (\text{an $\fo$ formula that does not mention any $X_i$}).
$$
Define $\varphi_i^{+}$, $\varphi_i^{-}$ and $\varphi_i^{\fo}$ as the formulas mentioned above.

In our procedure we assume that in $\varphi_i^{+}\wedge\varphi_i^{-}$ every first-order variable from $\bar{x},\bar{y}$ is mentioned at most once. If not, we add new first-order variables from $\fv$ to $\bar{y}$ so that no variable is repeated, replace them in $\varphi_i^{+}\wedge\varphi_i^{-}$ and add their respective equalities in $\varphi_i^{\fo}$ so that the new formula is equivalent. For example, if $\bar{x} = x$, $\bar{y} = y$ and $\varphi_i = X(x,y)\wedge \neg X(x,x) \wedge x < y$, then we redefine $\bar{y} = (x_1,x_2,x_3,x_4,y)$ and $\varphi_i := X(x_1,x_2) \wedge \neg X(x_3,x_4) \wedge x < y \wedge x_1 = x \wedge x_2 = y \wedge x_3 = x \wedge x_4 = x.$

Furthermore, we assume that $\varphi_i^{\fo}$ defines an ordered partition of the variables in the tuple $(\bar{x},\bar{y})$. For example, let $\bar{x} = (x_1,x_2,x_3,x_4)$. A possible ordered partition would be defined by the formula $\theta(\bar{x}) = x_2 < x_1 \wedge x_1 = x_4 \wedge x_4 < x_3$. On the other hand, the formula $\theta'(\bar{x}) = x_1 < x_2 \wedge x_1 < x_4 \wedge x_2 = x_3$ does not define an ordered partition since both $\{x_1\}<\{x_2,x_3\}<\{x_4\}$ and $\{x_1\} < \{x_2,x_3,x_4\}$ satisfy $\theta'$.
For a given $k$, we define $\cB_k$ as the number of possible ordered partitions for a set of size $k$. 
If $\varphi_i^{\fo}$ does not define an ordered partition, then for $1 \leq j \leq \cB_{\length(\bar{x},\bar{y})}$ 
let $\theta^j(\bar{x},\bar{y})$ be a formula that defines an ordered partition over $(\bar{x},\bar{y})$. The formula $\varphi_i(\bar{x},\bar{X},\bar{y})$ is converted into $\bigvee_{j = 1}^{\cB_{\length(\bar{x},\bar{y})}}\varphi_i(\bar{x},\bar{X},\bar{y}) \wedge \theta^j(\bar{x},\bar{y})$. Consider each of these disjuncts as a new $\varphi_i$.

Now let $\bar{u}$ be the tuple of every first-order variable mentioned in $\varphi_i^{+}$ and let $\bar{v}$ be such that $(\bar{x},\bar{y}) = (\bar{u},\bar{v})$. We define a formula $\mu_i$ that is satisfied only by the lexicographically smallest $\bar{u}$ that satisfies $\varphi_i^{+}$ for some $\bar{v}$:
$$
\mu_i(\bar{u}) = \exists\bar{v}\,\varphi_i^{\fo}(\bar{u},\bar{v})\wedge\forall\bar{u}'\forall\bar{v}'(\varphi_i^{\fo}(\bar{u}',\bar{v}')\to\bar{u}\leq\bar{u}').
$$
Lastly, we define $\kappa(\varphi_i)$:
$$
\kappa(\varphi_i) := 
$$
%Consider an $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$ where $\z = (z_1,\dots,z_d)$ and $\X = (X_1,\dots,X_r)$. There are three possibilities regarding the size of the tuples of free variables $\z$ and $\X$: (1) $d>0$ and $r=0$ (2) $d=0$ and $r>0$ (3) $d,r>0$. This separates the proof in three cases:
%	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$. That is,
%	\begin{eqnarray*}
%		f(\A) &=& \mid \{ \langle\e\rangle \mid \A \models \exists \x \ \varphi(\x,\e) \} \mid,
%	\end{eqnarray*}
%	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\e \in A^d$. Our goal here is to eliminate the lexicographically smallest sequence of variables, which can be done easily. First, let $\y = (y_1,\dots,y_k)$, $\y\,^\prime = (y_1^\prime,\dots,y_k^\prime)$ and
%	\begin{eqnarray*}
%		\varphi_{k,<}(\y\,^\prime,\y) &=& \bigvee_{i = 1}^k \left( \bigwedge_{j=1}^{i-1} y_j^\prime = y_j \wedge y_i^\prime < y_i \right).
%	\end{eqnarray*}
%	This formula is true if $\y\,^\prime$ is lexicographically smaller than $\y$. Now, let $f'$ be defined by
%	\begin{eqnarray*}
%		\varphi^\prime(\x,\z) &=& \varphi(\x,\z) \wedge \exists \z\,^\prime (\varphi(\x,\z\,^\prime) \wedge \varphi_{d,<}(\z\,^\prime,\z ) ).
%	\end{eqnarray*}
%	If $f(\A)>0$, then $f'(\A)$ will count exactly one element less than $f(\A)$. Otherwise, if $f(\A)=0$, then $\A \not\models\exists \x\,\varphi(\x,\e)$ for every tuple $\e$ of elements in $A$, so $\A \not\models\exists \x\,\varphi^\prime(\x,\e)$ for every $\e$ and, therefore, $f'(\A)=0$. Hence, $f' = f\dotminus 1,$ from which we conclude that $f\dotminus 1\in\E{1}.$
%	
	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,\dots,x_d)$ and $\X = (X_1,\dots,X_r)$. That is,
	\begin{eqnarray*}
		f(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \exists \x \ \varphi(\x,\P) \} \mid \label{f1},
	\end{eqnarray*}
	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\P = (P_1,\ldots,P_r)$ and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$. For the time being, suppose that
	\begin{eqnarray}
	\varphi(\x,\X) &=& \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)
	\end{eqnarray}
	where $n$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, $\y$ is a $p$-tuple of variables in $\x$, $\varphi^{-}(\X,\y)$ is a conjunction of negated predicates in $\X$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\beta(\x)$ is an FO-formula over $\L$ which mentions all variables in $\x$. Note that $\theta(\x)$ also mentions all variables in $\x$. We also assume that $(\x_1,\dots,\x_n,\y) = \x$. As an example, the following formula is of this form:
	\varphi(\x,\X) =  X_1(x_1,x_2) \wedge X_3(x_3) \wedge X_2(x_4,x_5) \wedge X_3(x_6) \wedge \neg X_1(x_7,x_8) \wedge \\ (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 \wedge x_2 = x_8 \wedge x_2 = x_5 \wedge x_8 < x_6 \wedge x_6 = x_7 ) \wedge \\ \forall z\big( S_1(x_1,z,x_2) \wedge x_3 = x_3 \wedge x_4 = x_4 \wedge x_5 = x_5 \wedge x_6 = x_6 \wedge x_7 = x_7 \wedge x_8 = x_8 \big),
	where $\x = (x_1,x_2,x_3,x_4,x_5,x_6,x_7,x_8)$ and $\X = (X_1,X_2,X_3)$. Here, $n = 4$, $\lambda(1) = 1$, $\lambda(2) = \lambda(4) = 3$ and $\lambda(3) = 2$, $\x_1 = (x_1,x_2)$, $\x_2 = (x_3)$, $\x_3 = (x_4,x_5)$, $\x_4 = (x_6)$ and $\y = (x_7,x_8)$. Moreover, $\varphi^{-}(\X,\y) = \neg X_1(x_7,x_8)$, $\theta(\x) = (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 \wedge x_2 = x_8 \wedge x_2 = x_5 \wedge x_8 < x_6 \wedge x_6 = x_7 )$, which defines a total order on the partition of $\x$ $\{\{x_1,x_3,x_4\},\{x_2,x_5,x_8\},\{x_6,x_7\}\}$, and $\beta(\x) = \forall z\big( S_1(x_1,z,x_2) \wedge x_3 = x_3 \wedge x_4 = x_4 \wedge x_5 = x_5 \wedge x_6 = x_6 \wedge x_7 = x_7 \wedge x_8 = x_8 \big)$.

@@

	\alpha_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha(\v_1,\dots,\v_n,\w)\wedge\bigvee_{i=1}^n(\u_i\neq\v_i)\big)\to (\u_1,\dots,\u_n)<(\v_1,\dots,\v_n)\Big)\Big],
